---
title: "modelling CPE - LAX"
author: "Henry"
date: '`r format(Sys.Date(), "%Y-%B-%d")`'
output: 
  html_document:
    highlight: zenburn
    theme: united
    df_print: paged
    toc: yes
    code_folding: hide
    code_download: true
    #fig_width: 17
    #fig_height: 7
---

<style>
slides > slide.backdrop {
  background: white;
}
</style>

<style>
.column-left{
  float: left;
  width: 76%;
  text-align: left;
}
.column-right{
  float: right;
  width: 24%;
  text-align: right;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# 0. Set up libraries
packages <- c("ggplot2", "tidyverse", "gridExtra", "ggthemes", "car", "radiant", "memisc")

for (p in packages) {
  if ((p %in% rownames(installed.packages())) == FALSE) {
    install.packages(p, dependencies = TRUE)
  }
  library(p, character.only = TRUE)
}
```


```{r}
carrier_names <- readRDS("~/git/airline-pricing/data/carrier_names.rds")
ticket_19 <- readRDS("~/git/airline-pricing/data/ticket_2019.rds") %>%
  mutate(
    Quarter = as.factor(Quarter),
    RoundTrip = as.factor(RoundTrip),
    OnLine = as.factor(OnLine),
    DistanceGroup = as.factor(DistanceGroup)
  )

# filter dollarCred = 1, and BulkFare = 0, only Southwest airlines only
data <- ticket_19 %>%
  filter(BulkFare == "0" & DollarCred == "1" & RPCarrier == "WN")

glimpse(data)
```


```{r}
dataWN <- ticket_19 %>%
  filter(BulkFare == "0" & DollarCred == "1" & RPCarrier %in% c("AA", "DL", "UA", "WN"))
```

```{r}
# Find out why?
max(dataWN$Passengers)
```

## Itinerary count by origin Airport ID

```{r}
by_airport <- dataWN %>%
  group_by(Origin) %>%
  summarise(countItin = n()) %>%
  ungroup() %>%
  mutate(percCount = countItin * 100 / sum(countItin)) %>%
  arrange(desc(countItin)) %>%
  slice(1:10)
```


```{r}
# slice(1:10)

by_airport %>%
  ggplot(aes(x = reorder(Origin, countItin), y = countItin)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  geom_text(aes(label = format_nr(countItin, dec = 0)), hjust = -0.3) +
  coord_flip()
```

```{r}
newData <- dataWN %>%
  filter(Origin %in% by_airport$Origin) %>%
  group_by(DistanceGroup, RPCarrier, Origin) %>%
  summarise(
    Demand = n(),
    Fare = mean(ItinFare)
  )

aa_dl <- newData %>%
  ungroup() %>%
  filter(RPCarrier %in% c("AA", "DL")) # %>%
# mutate(id = rep(1:40, 2))
```

```{r}
# data check
dataWN %>%
  filter(RPCarrier == "WN")
```


* Adding DistanceGroup alone isn't suggestive as it takes up the variance.

```{r}
# Function to get a specific airline's RoundTrip (1/0) data by Origin (Top 10)
data_munge <- function(df, RT, Carrier) {
  newData <- df %>%
    filter(Origin == 'LAX') %>%
    group_by(DistanceGroup, RPCarrier, RoundTrip, Origin) %>%
    summarise(
      Demand = n(),
      Fare = mean(ItinFare)
    )

  aa_dl <- newData %>%
    ungroup() %>%
    filter(RoundTrip == RT & RPCarrier %in% c(Carrier)) %>%
    mutate(id = paste0(DistanceGroup, RoundTrip, Origin))

  names(aa_dl)[names(aa_dl) == "Demand"] <- paste0("RPCarrierDemand", Carrier)
  names(aa_dl)[names(aa_dl) == "Fare"] <- paste0("RPCarrierFare", Carrier)


  return(aa_dl)
}
```


```{r}
# sample run
data_munge(dataWN, 1, "AA")
```

```{r}
dfAA <- rbind(
  data_munge(dataWN, 1, "AA"),
  data_munge(dataWN, 0, "AA")
)


dfDL <- rbind(
  data_munge(dataWN, 1, "DL"),
  data_munge(dataWN, 0, "DL")
)

dfUA <- rbind(
  data_munge(dataWN, 1, "UA"),
  data_munge(dataWN, 0, "UA")
)

dfWN <- rbind(
  data_munge(dataWN, 1, "WN"),
  data_munge(dataWN, 0, "WN")
)

dfFSC <- dfWN %>%
  left_join(dfDL %>% dplyr::select(id, RPCarrierDemandDL, RPCarrierFareDL), by = c("id")) %>%
  left_join(dfUA %>% dplyr::select(id, RPCarrierDemandUA, RPCarrierFareUA), by = c("id")) %>%
  left_join(dfAA %>% dplyr::select(id, RPCarrierDemandAA, RPCarrierFareAA), by = c("id"))
```

```{r}
hist(dfFSC$RPCarrierDemandAA)
hist(log(dfFSC$RPCarrierDemandAA))
hist(dfFSC$RPCarrierFareAA)
hist(log(dfFSC$RPCarrierFareAA))
```


## WN is the subject


```{r}
# data transformation
dfFSC <- drop_na(dfFSC)

nrow(dfFSC)
farecols <- colnames(dfFSC)[(grepl("Fare", colnames(dfFSC)))]


dfFSC[, farecols][dfFSC[, farecols] == 0] <- 1


lm.1 <- lm(log(RPCarrierDemandWN) ~ log(RPCarrierFareWN) + RoundTrip, data = dfFSC)
summary(lm.1)
```

```{r}
# data transformation
dfFSC <- drop_na(dfFSC)

nrow(dfFSC)
farecols <- colnames(dfFSC)[(grepl("Fare", colnames(dfFSC)))]


dfFSC[, farecols][dfFSC[, farecols] == 0] <- 1


lm.2 <- lm(log(RPCarrierDemandWN) ~ log(RPCarrierFareWN) + log(RPCarrierFareUA) + log(RPCarrierFareAA) + log(RPCarrierFareDL) +  RoundTrip, data = dfFSC)
summary(lm.2)
```

```{r}
vif(lm.2)
```

```{r}
# data transformation
dfFSC <- drop_na(dfFSC)

nrow(dfFSC)
farecols <- colnames(dfFSC)[(grepl("Fare", colnames(dfFSC)))]


dfFSC[, farecols][dfFSC[, farecols] == 0] <- 1


lm.3 <- lm(log(RPCarrierDemandWN) ~ log(RPCarrierFareWN) + log(RPCarrierFareUA) * log(RPCarrierFareAA) + log(RPCarrierFareDL) + RoundTrip, data = dfFSC)
summary(lm.3)
```

```{r}
vif(lm.3)
```

```{r}
# data transformation
dfFSC <- drop_na(dfFSC)

nrow(dfFSC)
farecols <- colnames(dfFSC)[(grepl("Fare", colnames(dfFSC)))]


dfFSC[, farecols][dfFSC[, farecols] == 0] <- 1


lm.4 <- lm(log(RPCarrierDemandWN) ~ log(RPCarrierFareWN) + log(RPCarrierFareUA) + log(RPCarrierFareAA) * log(RPCarrierFareDL) + RoundTrip, data = dfFSC)
summary(lm.4)
```

* AA is -ve and DL is +ve

```{r}
vif(lm.4)
```

```{r}
# data transformation
dfFSC <- drop_na(dfFSC)

nrow(dfFSC)
farecols <- colnames(dfFSC)[(grepl("Fare", colnames(dfFSC)))]


dfFSC[, farecols][dfFSC[, farecols] == 0] <- 1


lm.5 <- lm(log(RPCarrierDemandWN) ~ log(RPCarrierFareWN) + log(RPCarrierFareAA) + log(RPCarrierFareUA) * log(RPCarrierFareDL) + RoundTrip, data = dfFSC)
summary(lm.5)
```

```{r}
mtable(lm.1, lm.2, lm.3, lm.4, lm.5)
```

## LCC

```{r}
dataLCC <- ticket_19 %>%
  filter(BulkFare == "0" & DollarCred == "1" & RPCarrier %in% c("B6", "WN", "F9", "G4", "NK"))
```

### Itinerary count by origin Airport ID

```{r}
by_airport <- dataLCC %>%
  group_by(Origin) %>%
  summarise(countItin = n()) %>%
  ungroup() %>%
  mutate(percCount = countItin * 100 / sum(countItin)) %>%
  arrange(desc(countItin)) %>%
  slice(1:10)
```


```{r}
# slice(1:10)

by_airport %>%
  ggplot(aes(x = reorder(Origin, countItin), y = countItin)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  geom_text(aes(label = format_nr(countItin, dec = 0)), hjust = -0.3) +
  coord_flip()
```


```{r}
### Building model data frame

dfB6 <- rbind(
  data_munge(dataLCC, 1, "B6"),
  data_munge(dataLCC, 0, "B6")
)


dfWN <- rbind(
  data_munge(dataLCC, 1, "WN"),
  data_munge(dataLCC, 0, "WN")
)

dfAS <- rbind(
  data_munge(dataLCC, 1, "F9"),
  data_munge(dataLCC, 0, "F9")
)

dfG4 <- rbind(
  data_munge(dataLCC, 1, "G4"),
  data_munge(dataLCC, 0, "G4")
)

dfNK <- rbind(
  data_munge(dataLCC, 1, "NK"),
  data_munge(dataLCC, 0, "NK")
)

dfLCC <- dfWN %>%
  left_join(dfB6 %>% dplyr::select(id, RPCarrierDemandB6, RPCarrierFareB6), by = c("id")) %>%
  left_join(dfAS %>% dplyr::select(id, RPCarrierDemandF9, RPCarrierFareF9), by = c("id")) %>%
  left_join(dfNK %>% dplyr::select(id, RPCarrierDemandNK, RPCarrierFareNK), by = c("id")) %>%
  left_join(dfG4 %>% dplyr::select(id, RPCarrierDemandG4, RPCarrierFareG4), by = c("id")) %>%
  left_join(dfDL %>% dplyr::select(id, RPCarrierDemandDL, RPCarrierFareDL), by = c("id")) %>%
  left_join(dfUA %>% dplyr::select(id, RPCarrierDemandUA, RPCarrierFareUA), by = c("id")) %>%
  left_join(dfAA %>% dplyr::select(id, RPCarrierDemandAA, RPCarrierFareAA), by = c("id"))
```

### Southwest is the subject

```{r}
lm.1 <- lm(log(RPCarrierDemandWN) ~ log(RPCarrierFareWN)  + RoundTrip, data = dfLCC)
summary(lm.1)
```


```{r}
lm.2 <- lm(log(RPCarrierDemandWN) ~ log(RPCarrierFareWN)  +  log(RPCarrierFareNK) +  RoundTrip, data = dfLCC)
summary(lm.2)
```

```{r}
lm.3 <- lm(log(RPCarrierDemandWN) ~ log(RPCarrierFareWN)  +  log(RPCarrierFareB6) + log(RPCarrierFareNK) +  RoundTrip, data = dfLCC)
summary(lm.3)
```

```{r}
vif(lm.3)
```


#### With NK, B6, F9

```{r}
lm.4 <- lm(log(RPCarrierDemandWN) ~ log(RPCarrierFareWN)  +  log(RPCarrierFareB6) + log(RPCarrierFareNK) + log(RPCarrierFareF9) + RoundTrip, data = dfLCC)
summary(lm.4)
```

```{r}
vif(lm.4)
```

```{r}
mtable(lm.1, lm.2, lm.3, lm.4)
```

### No log-log 

```{r}
lm.WN <- lm((RPCarrierDemandWN) ~ (RPCarrierFareWN)  +  (RPCarrierFareB6) + (RPCarrierFareNK) + (RPCarrierFareF9) + 
              (RPCarrierFareAA) + RoundTrip, data = dfLCC)
summary(lm.WN)

```

### Models for price setting with many competitors' price included


```{r}
lm.WN <- lm(log(RPCarrierDemandWN) ~ log(RPCarrierFareWN)  +  log(RPCarrierFareB6) + log(RPCarrierFareNK) + log(RPCarrierFareF9) + 
              log(RPCarrierFareAA) + RoundTrip, data = dfLCC)
summary(lm.WN)

```




### B6 is the subject

```{r}
lm.B6 <- lm(log(RPCarrierDemandB6) ~ log(RPCarrierFareWN)  +  log(RPCarrierFareB6) + log(RPCarrierFareNK) + log(RPCarrierFareF9) + RoundTrip, data = dfLCC)
summary(lm.B6)

```

```{r}
lm.F9 <- lm(log(RPCarrierDemandF9) ~ log(RPCarrierFareWN)  +  log(RPCarrierFareB6) + log(RPCarrierFareNK) + log(RPCarrierFareF9) + RoundTrip, data = dfLCC)
summary(lm.F9)

```

```{r}
mtable(lm.B6, lm.F9, lm.WN)
```

```{r}
# nrow(dfLCC)
# # data transformation
# df <- drop_na(dfLCC)
# 
# nrow(dfLCC)
# farecols <- colnames(df)[(grepl("Fare", colnames(dfLCC)))]
# 
# 
# dfLCC[, farecols][dfLCC[, farecols] == 0] <- 1
# 
# lm.CP <- lm(log(RPCarrierDemandWN) ~ log(RPCarrierFareWN)  +  log(RPCarrierFareB6) + log(RPCarrierFareF9) + log(RPCarrierFareG4) , data = dfLCC)
# summary(lm.CP)
```

* Data is too sparse and missing
